### 2017-05-16 doc/response.tex
paper files added

\documentclass{article}
\usepackage{a4wide}
\usepackage{xr-hyper} 
\usepackage{hyperref}
\externaldocument{paper}
\title{Response to the Reviewers}
%\author{Ákos Kádár \and Grzegorz Chrupala \and Afra Alishahi}
\author{Submission number: 882 \and
Action editor:  Christopher Potts }
\date{}
\begin{document}

\maketitle
The Editor's and reviewers' comments were of great help and enabled us
to make important improvements on our previous submission. Some minor or major
changes were made to almost all of the sections of the paper.
One of the major changes in the revised version is the addition of
discussions on the generality and applicability of the proposed
methods in sections \ref{sec:intro}, \ref{sec:computeomission}, 
\ref{sec:beyondlexical} and \ref{sec:reprdim}. The revised version
still focuses on the {\sc Imaginet} architecture, but in addition 
allocates more space and more emphasis on the discussion
of the introduced analysis techniques.
The other major revision concerns the re-structuring 
of Section \ref{sec:macro}. In the revised version we 
separated the description of the omission score computation 
(Section \ref{sec:computeomission}) and
the discussion regarding the distribution of the omission scores over 
categories of words for the different models 
(Section \ref{sec:omitimaginet}). 
Furthermore, in Section \ref{sec:beyondlexical}
we introduced a common methodology
to assess the impact of position and grammatical
function of tokens (while controlling for word-types) based on the 
omission scores. We also provide finer-grained analysis
of these factors separately in sections \ref{sec:gramfunc} and
\ref{subsec:information-struct}. Thanks to the reviewers' comments
we realized that the original submission gave the impression that our main
result was the amount of attention
different models pay to words of different categories.  We hope that the
revised version helps the reader to also focus on
sections \ref{sec:beyondlexical}, \ref{sec:gramfunc} and
\ref{subsec:information-struct} which explore linguistic features
that the networks encode on top of simple lexical features. These and
the additional less substantial revisions are discussed in more
detail below.

\begin{verbatim}

>> Editor's Comments <<

The paper is pitched in very broad terms (methods for analyzing RNNs), 
but the methods are actually quite specific to classifier-type RNNs, 
and many aspects of the work seem applicable only to the Imaginet 
setting. I would suggest clarifying the scope of the claims. Similarly, 
as Reviewer A points out, some of the rhetoric is misaligned with the 
methods and results.  
\end{verbatim}

This point was raised by both \emph{Reviewer A} and \emph{Reviewer B} and 
had a major impact on our paper.
Sections \ref{sec:intro}, \ref{sec:computeomission}, 
\ref{sec:beyondlexical} and \ref{sec:reprdim} include new paragraphs
aimed to make it clear in which settings the described methods could be meaningfully applied.

\begin{itemize}
  \item In Section \ref{sec:intro} ``Introduction'' on page
    \pageref{generalintro}, we point out that we applied the introduced methods
   only to {\sc Imaginet}, 
  but that we also discuss their generalizability and offer pointers
  on how to apply them in other settings.
  \item On page \pageref{edit:introomissiongeneral} after briefly introducing the 
   \emph{omission score} technique we point out that it can be applied to
   other architectures too. The next paragraph on page
   \pageref{edit:introtopkgeneral}, after introducing the
   \emph{top-K-contexts} method itself, discusses its
  applicability to other settings. 
  \item We added a paragraph to the end of Section \ref{sec:computeomission} on
  page \pageref{edit:omitgeneral} that points out in a bit more detail 
  how the \emph{omission} technique can be generalized to other
  RNN variants, recursive neural networks and CNNs. 
  We hope that it will be clear to the reader that the analysis 
  provided in Section \ref{sec:omitimaginet} can be applied to the 
  same models as the \emph{omission} method.
  \item We also added text in the first paragraph of Section \ref{sec:beyondlexical}
  on page \pageref{edit:beyonlexicalgeneral} to clarify in which settings 
  the described methods could be meaningfully applied.
  \item In Section \ref{sec:reprdim} we explain that the provided 
  analysis can be straight-forwardly applied to other RNN variants and 
  suggest how to apply it to multi-layer architectures. We also point
  out that it is likely to be limited to uni-directional models.
\end{itemize}

\begin{verbatim}
The core results for content words and functions 
words seem like the expected ones given their 
linguistic content. In particular, the omission
score method seems to show that the networks learn 
to identify discriminating content words, and that
changing or removing those words has a dramatic effect 
on final predictions. This seems like a prerequisite 
(given the linguistic facts) for learning an effective classifier. 
Regular train/dev/test experimental analyses already
show that these models are effective. Could you further 
articulate what we learn via the omission score method 
that we don't get from standard evaluations? 
\end{verbatim}

As mentioned above, the fact that the {\sc Visual} network learns to
identify discriminating content words is only one of several results
in our paper. Even this basic finding is not as obvious as it
may seem: in the revised version, the last paragraph in Section \ref{sec:omitimaginet} on page 
\pageref{edit:generality} discusses that the relatively low omission scores
for verbs for the {\sc Visual} pathway are somewhat unexpected as one 
would expect high omission
scores for content words in general. We go on to discuss that the caption to
image prediction setting 
seems to lead to more task specific representations than expected.

Furthermore, much of the paper is devoted to investigating patterns
over and above attention paid to particular lexical items: for
example in our experiment in Section~\ref{sec:macro} we focus on the
effect of grammatical role and position in the sentence, {\it while
  controling for word-type.} (See also the comment below).

We believe these insights go well beyond what standard dev/test evaluations
can tell us.

\begin{verbatim}

The connections with POS tags and dependency relations are intriguing.
Could you, though, clarify the significance of these findings? My concern is
that they just reflect the fact that the content-word/function-word
distinction is highly correlated with POS and dependency relation.
\end{verbatim}

This comment really helped us restructure Section \ref{sec:macro}, with the aim
to describe and highlight our main findings in a more clear manner. 
Firstly in the new version of Section \ref{sec:omitimaginet} we do not emphasize so much
the content- vs. function-word distinction and point out the relatively
low omission of verbs in case of {\sc Visual}. Furthermore, in the new version
we dedicate Section \ref{sec:beyondlexical} to raise the question about 
the correlation between word-types, dependency relations and position in the 
sentence and describe how we disentangle these factors using linear
regression modeling. Concretely, we show
that the omission scores both in case of {\sc Visual} and {\sc Textual} do not
only depend on the word types, but on the grammatical function 
and position of the tokens, while controling for word types.
To further illustrate this point, we introduce a non-sequential baseline model {\sc Sum}, 
in which we replace the RNN with a
vector-summation operation over learned word-embeddings.

\begin{verbatim}
I had trouble understanding the information structure section. 
I agree with the core insight about old/topic before new/comment 
in languages like English. I had trouble relating this to what 
I would expect from the RNN. For instance, the structure of the 
RNN seems to ensure that information nearer the end of the sequence
will have a larger impact. Is the Textual omission score result due 
to this factor or is it truly due to something about language? 
Conversely, for Visual, I guess I would have anticipated exactly 
the opposite of what is reported here: the comment part should
be the most important, corresponding to the new and visually 
salient elements, rather than those that can fade into the background.
\end{verbatim}

Motivated by this and the previous comment in the revised version
sections \ref{sec:gramfunc} and \ref{subsec:information-struct} use the same methodology
for the analysis of the significance of grammatical function and position in the sentence.
We hope this modification makes our methodology and the interpretation of our 
results more straight-forward. Also for clarity, we streamlined the visualization
of comparing how the position of tokens affect the omission scores of the different models
- Figure \ref{fig:posrqs} on page \pageref{fig:posrqs}. 
The current plot shows the same effect for {\sc Textual}
and similar for {\sc Visual} as before, but hopefully in a more interpretable
way. Furthermore, we have added a short discussion 
on topic vs. comment in the image-description setting on page 
\pageref{edit:topiccomment}, which we agree was definitely needed.

\begin{verbatim}

>> Reviewer A <<

One question about the omission score: the lower the score is, the low the
similarity is. Thus, when dropping the important words, the similarity
should go down -- so a lower omission score. It is not clear to me why
dropping important words will lead to a higher omission score.

\end{verbatim}

Thank you for pointing this mistake out, we actually
compute the cosine distance $1-cos(x,y)$ not cosine.
We modified Equation \ref{eg:omit} on page \pageref{eg:omit} accordingly.

\begin{verbatim}
First, why is the architecture of
Chrupala et al (2015) selected only? Is it representative enough for general
RNNs? What about LSTM and its variants for language modelling task? How
would they be different? It would be good to see such discussions. 
\end{verbatim}

Section \ref{sec:intro} (``Introduction'')
discusses the reason why we choose this particular architecture
on page \pageref{explainimaginet}. The model contains 
two separate RNNs: a language model and a
model that predicts image-vectors given sentences. 
These ``pathways'' are trained on the same training data and 
share word-embeddings. We choose such a setup as we
are interested in 1) analyzing the hidden activation patterns 
(rather than word embeddings), and 2) developing methods that allow for comparing models.
As described in the first answer to the 
Editor's first question, not discussing the generality and
applicability of our methods was definitely a shortcoming of the paper and
in the new version sections \ref{sec:computeomission},
\ref{sec:beyondlexical}, \ref{sec:reprdim} discuss
how general our techniques are and how to apply them in new settings.

\begin{verbatim}
Second, most of the paper presents qualitative results, without rigorous
quantitative justifications. It would be nice, however, if the authors could
collect some manual annotations and conduct automatic evaluations based on
such annotations. For example, such annotations could include human's
judgements on how important each word (or word class, syntactic class) in
the sentence are etc.
\end{verbatim}
Regarding the collection of manual annotations, this is a great
suggestion and we also have been planning 
to assess the relationship between our results and human 
performance. Conducting such experiments with 
reliable results is labor- and resource-intensive and for the time 
being we consider such an experiment to be out of the scope of our current
paper, but as a great direction for our future work. 
We discuss this idea in the new version of Section 
\ref{conclusion} ``Conclusion''.

Regarding qualitative results, we believe the paper offers a good
balance of quantitative and qualitative analyses, even though our
quantitative results do not take the form of accuracy scores. Our
quantitative results include:
\begin{itemize}
 \item We introduce omission scores in Section \ref{sec:computeomission}
 with the aim to quantify the importance of words to 
 the final prediction.
 \item Section \ref{sec:omitimaginet} compares {\sc Visual} and {\sc Textual} 
 with respect to the difference between their omission score distributions
 over POS tags and deprel categories.
 \item The new Section \ref{sec:beyondlexical} fits linear models with 
 various sets of predictor variables to predict omission score as outcome and
 assesses the contribution of these variables. 
 The analysis in sections \ref{sec:gramfunc}, 
 \ref{subsec:information-struct} also utilize these models.
 \item In Section \ref{seq:describe} we compare 
 {\sc Visual} and {\sc Textual}
 by measuring the strength of association between the 
 binned activation values
 of the hidden units and the token/deprel n-grams in the corpus with mutual
 information. 
\end{itemize}

\begin{verbatim}
Finally, although the visualisation is helpful 
and it might help the readers understand the 
underlying activities of the RNN, little is known about why. 
I understand it is just an empirical paper, however, it would 
be helpful to at least share with the readers some high level 
insights so that the readers could learn more from this paper. 
One question that I have in mind after reading this paper is: 
these results related to RNN presented here are expected. 
It's good that RNN is able to capture such things. 
However, how do I make use of the results here in my research? 
How does this have some impact on my work?
\end{verbatim}

RNN variants proved to be powerful tools for modeling various aspects of human
language and for NLP applications, and yet there has been little work on showing 
what these models learn about language. In our paper we explore the 
learned representations from a linguistic point of view, in the spirit
of basic research.  
Engineering applications are not the focus of this paper, and so in
the body of the paper we try not to speculate on the relevance of our
findings to practical NLP.

That said, using our omission score results
we do observe that the task of predicting images 
from descriptions on MS-COCO does
not promote the learning of representations for verbs. 
Furthermore, one can observe a recency
effect in the meaning representation of {\sc Textual}, 
which probably also leads to too task specific representations. 
Thus, we do hope that our methods can provide help for such
engineering research in the future.

\begin{verbatim}
The authors claimed at the end of the paper that 
"The goal of our paper is to propose novel methods
for the encoding of linguistic knowledge in RNNs" -- I am not sure if I am
convinced that this is goal of this work.  It is largely an empirical paper
that tries to visualise/understand the RNN which was previously regarded as
a blackbox, but does not really "propose" methods for encoding linguistic
knowledge in RNN.
\end{verbatim}

This was an unfortunate mistake to leave this typo in the final version;
we meant to write ``The goal of our paper is to propose novel methods
for the ANALYSIS of the encoding of linguistic knowledge in RNNs''. We fixed it and 
thank you for noticing it!

\begin{verbatim}
 >> Reviewer B <<

The main criticism I would have of this paper is that it 
sets out as a providing generic new method for analysing RNNs 
with text input, amongst other modalities. 
I think this is overstating the contents of the paper: 
the methods of analysis are fairly tied to the nature of the model
(unidirectional RNN without attention) and broad task 
(producing a final representation that can be used for classification or generation). 
It would have been interesting to the same analyses performed on other, 
related models (bi-directional RNN, word-based log-bilinear models, etc), 
and on RNN tasks with other objectives (machine translation, language modelling, 
sequence labelling). I realise it's a tall order to add this to the current work,
but perhaps re-stating the scope of the work 
(providing an example of how to perform further qualitative 
and quantitative analysis of a an existing model trained
 on a specific task) 
 and briefly discussing how to adapt the experiments to other domains would be good.
\end{verbatim}

We really appreciated this comment! As described in our first answer, 
this observation had a large impact on our paper. We did add 
discussions in sections \ref{sec:intro}, \ref{sec:computeomission} 
,\ref{sec:beyondlexical}, \ref{sec:reprdim} on how to apply our 
introduced methods to different types of models.
We hope the Reviewer agrees that the changes we made based on 
this comment made our paper much stronger.

\begin{verbatim}
Likewise, a discussion comparing and contrasting these methods
of analysis with other (complementary?) visualisation methods 
for RNNs with attention such at Badhanau et al. 2015 or 
Rocktaschel et al. 2015 [ICLR 2016] (which could be cited here) would be welcome.
\end{verbatim}

This was also a great suggestion and we did include a 
discussion of the attention-weight 
visualization on page \pageref{edit:attention} in 
Section \ref{sec:related} based on the suggested works
``Related Work''. This has made the section much more complete.

\begin{verbatim}
It would be nice to get a clear statement of the objective being optimised
against in eq 9.
\end{verbatim}

The two loss functions are now described by equations \ref{eq:lossce}, and 
\ref{eq:losscos} and the multi-task loss by Equation \ref{eq:losscombo}.

\begin{verbatim}
Figure 2 a little hard to read.
\end{verbatim}

We were considering ways to make the metioned figure less cluttered, but
we are already reducing clutter by
only including labels with frequency at least 500. We feel increasing
this threshold further or merging categories would risk 
removing too much information and end up with a misleading picture. 
To help the understanding of the figure the main trends are explained in the text.

\begin{verbatim}
Figure 6 very hard to read.
\end{verbatim}

We agree that both the figure and the methodology in the previous
version concerning the models' sensitivity to position in
sentences was  not easy to interpret. Section 
\ref{subsec:information-struct} reports new results 
regarding the significance of the position of tokens along
with a less cluttered Figure \ref{fig:posrqs}.

\begin{verbatim}
Section 5.1 "textual is more sensitive to ..." -> it looks like textual is
more sensitive to most words being removed to some extent. I don't know if
it's fair to single out these categories. Instead, just contrast more
clearly with the heavy focus on content words (mainly NNs) for Visual.
\end{verbatim}

For clarity in the revised version of the paper we separated the
introduction of the omission scores and the results based on them. 
Section \ref{sec:computeomission} motivates omission scores and describes
how to compute them and Section \ref{sec:omitimaginet} describes the
distribution of the omission scores over POS-tags and deprels for 
{\sc Textual} and {\sc Visual}. On page \ref{edit:textualomission} 
in Section \ref{sec:omitimaginet} we point out that {\sc Visual} is 
only sensitive to a couple of word-types, 
while {\sc Textual} has a flatter omission score distribution in general.
Only after this observation we go on to single out a couple of
categories for a more detailed comparison.

\begin{verbatim}
Section 5.2, model 1 vs model 2 <- is the a significant difference in
number of features here? Model 2 seems like it would have more features, and
intuitively, more info == better model (modulo feature quality), so quick
comment on this here would be good.
\end{verbatim}

We re-structured the sections analyzing the impact of position 
and grammatical function. In the new version sections
\ref{sec:gramfunc} and \ref{subsec:information-struct} 
apply the same methodology to explore both factors. 
Instead of commenting  on the difference in the number of features, in the new
version for the comparison of {\sc Visual} and {\sc Textual} 
we introduce a non-sequential baseline model - {\sc Sum}, and apply
the same regression models with the same features to this baseline.
Adding position and deprel variables in the linear model
to predict omission scores does slightly increase the $R^2$ 
even in case of {\sc Sum} - probably
due to unseen words -, but much more so for {\sc Visual} and {\sc
  Textual}.

\begin{verbatim}
Section 5.3 is very tied to structure of the model use in this paper
(would expect different results for bidirection RNN, recursive neural net).
Comment on this here would be good.
\end{verbatim}

The first paragraph in section Section \ref{sec:beyondlexical}
on page \pageref{edit:beyonlexicalgeneral} argues that the analysis
presented regarding the significance of word position are tied to RNNs. 
We also added a footnote on page \pageref{edit:foot} 
pointing out that recursive neural networks
and CNNs are not affected by left-to-right and right-to-left word order
in the way RNNs and bi-RNNs are. We agree that the results 
would be probably different for bidirectional RNNs, but these 
architectures are still affected by word-order.

\begin{verbatim}
Not a fan of qualitative analysis as done in Table 1/2. Too easy
over-interpret, but I realise this is personal taste. Would not be offended
if this is left in, but would suggest scrapping this to create space for
discussions suggested above.
\end{verbatim}

Both in the Section \ref{sec:intro} and Section \ref{sec:reprdim}
we point out that the provided results are based on qualitative methods.
We agree that Tables \ref{tab:contexts}, \ref{tab:syntax} 
and also Figure \ref{fig:dimheat} are easy 
to over-interpret. This is arguably a danger with 
qualitative analysis in general, but we do believe that it is interesting
to complement our quantitative analysis with a detailed exploration
of the function of individual hidden units from a closer look.

\begin{verbatim}
 >> Reviewer C <<

The paper builds heavily on previous work "IMAGINET", is densely described
but lacking details in certain sections. For example, I find it rather
challenging to understand the following:

- the process of image vector prediction is vague (eqn 7). Does it mean
prediction at the pixel level of an entire image, or a blob or principal
image segment of interest? Also, how does the image retrieval shown in
Figure 1 operate?

\end{verbatim}

To make these points more clear:

\begin{itemize}
	\item On page \pageref{edit:dumdumeddy} we 
added a footnote ``Representing the full image, extracted from 
a pre-trained Convolutional Neural Network''.
	\item On page \pageref{edit:retrievalexplain} in Section 
	\ref{sec:computeomission} we added an explanation about the retrieval
	process.
\end{itemize}

\begin{verbatim}
- Figure 6 would benefit from showing the aggregated POS groupings such as
"VBN,VBG,VB" all grouped into "VERB" etc. for benefit of clarity in a small
graph (too cluttered)
\end{verbatim}

Thank you for pointing it out, we hope that the reviewer agrees that 
the new Figure \ref{fig:posrqs} is less cluttered and more intuitive.

\end{document}