### 2017-05-16 doc/conclusions.tex

\section{Conclusion}
\label{conclusion}

The goal of our paper is to propose novel methods for the 
analysis of the encoding of linguistic knowledge in RNNs trained on language tasks.
We focused on developing quantitative methods to measure the importance
of different kinds of words for the performance of such models. Furthermore, we
proposed techniques to explore what kinds of linguistic features the models
learn to exploit on top of lexical cues.

Using the {\sc Imaginet} model as our case study,
our analyses of the hidden activation patterns show that the {\sc Visual} model
learns an abstract representation of the information structure of the
language, and pays selective attention to lexical categories and
grammatical functions that carry semantic information. In contrast,
the language model {\sc Textual} is sensitive to features of a more
syntactic nature. We have also shown that each network contains
specialized units which are tuned to both lexical and structural
patterns that are useful for the task at hand, some of which can carry
activations to later time steps to encode long-term dependencies.

In future we would like to apply the techniques introduced in this paper 
to analyze the encoding of linguistic form and function of 
recurrent neural models trained on different objectives, 
such as neural machine translation systems
\cite{sutskever2014sequence} or the purely distributional
sentence embedding system of \cite{kiros2015skip}. A number of
recurrent neural models rely on a so called attention mechanism, first
introduced by \namecite{bahdanau2014neural} under the name of soft
alignment. In these networks attention is explicitly represented and it
would be interesting to see how our method of discovering implicit
attention, the omission score, compares. For future work we also propose to
collect data where humans assess the importance of each word in sentences
and explore the relationsship between \emph{omission scores} for various models
and human annotations.\label{edit:humanjudgement} Finally, one of the benefits
of understanding how linguistic form and function is represented in RNNs is that
it can provide insight into how to improve systems. We plan to draw on
lessons learned from our analyses in order to develop models with better 
general-purpose sentence representations.

% \iffalse
% We developed \emph{macro} and \emph{micro} level methods to analyze the 
% activation patterns of Recurrent Neural Networks
% from a linguistic point of view. On the \emph{macro} level
% we introduced the $\mathrm{omission}$ score
% to measure the salience of tokens in sentences and
% showed how aggregating these scores in terms of part-of-speech categories and
% grammatical functions allows for a more in-depth understanding of the kinds of
% linguistic structure RNNs learn from linguistic data. In particular we have shown
% that {\sc Visual} learns to pay attention to tokens depending of their grammatical
% function. In these experiments we focused on how the model interprets
% the same \emph{nouns} fulfilling different grammatical functions, but in future work 
% this method can be straight-forwardly applied to other classes of content-words 
% such as \emph{verbs} or \emph{adjectives}. In addition we provided evidence 
% that {\sc Visual} pays more attention to nouns that are closer to 
% the beginning of the sentence, which is motivated by the information structure of English.
% In our \emph{micro} level analyses we developed a simple and general purpose method
% dubbed \emph{top K contexts}, which describes the function of a particular hidden unit 
% by ranking the contexts that produce the highest activation values for that unit.
% We performed exploratory analysis on these contexts, especially focusing on 
% hidden units whose activations are predictive of the grammatical function of tokens.
% We observed that in numerous cases units are highly activated for contexts that represent
% combined syntactic semantic templates. Furthermore, we explored dimensions that
% carry over their activations to represent longer term dependencies and demonstrated 
% a visualization technique to explore these patterns. Lastly, we performed a comparison
% between {\sc Textual} and {\sc Visual} using the mutual information between the activation
% values of their hidden units and the contexts. Our analysis showed a high level difference
% between the kinds of features the two pathways extract, in that the
% linguistic regularities encoded by hidden units of {\sc Textual} 
% seem to be more characterized by syntactic templates than in case of {\sc Visual}.
% In addition to the insights provided by the methods implemented in this paper
% we believe them to be general enough to allow  cognitive linguistics research 
% to further explore the linguistic knowledge in the activation patterns
% of RNNs trained on large scale data sets and on challenging tasks.

% \fi