### 2017-05-16 doc/omission-stat/analysis.Rmd

---
output: html_document
---
## Analysis of omission scores for textual and visual pathway of Imaginet

```{r, echo=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
read.data <- function(f) { 
  data <- read.csv(f, col.names= c('omit','id','word','pos','deprel', 'score')) %>% 
  select(-omit) %>% na.omit()
  return(data[! data[,'pos'] %in%c('<END>','END'),])
}
data_v <- read.data('../data/visual_full_omission.csv')
data_v$pathway <- "visual"
data_v$deprel <- reorder(data_v$deprel, data_v$score, median)
data_v$pos <- reorder(data_v$pos, data_v$score, median)
data_t <- read.data('../data/textual_full_omission.csv')
data_t$pathway <- "textual"
data_t$deprel <- reorder(data_t$deprel, data_v$score, median)
data_t$pos <- reorder(data_t$pos, data_v$score, median)

data <- rbind(data_t , data_v) %>% mutate(pathway=as.factor(pathway))
```
```{r}
filter_by_count <- function(dat, col, count) {
  dat[dat[,col] %in% names(table(dat[,col]))[table(dat[,col]) > count],]
}

MINCOUNT<-500
subset <- filter_by_count(data %>% filter(pathway=='visual'),"pos", MINCOUNT)
ggplot(subset, aes(x=pos, y=score)) + 
  geom_boxplot(outlier.colour = 'grey', outlier.size = 1, varwidth = FALSE) +
  coord_flip() +
  theme(text=element_text(size=15))
ggsave(file="visual-omission-pos-boxplot.png",width=5, height=5)

subset <- filter_by_count(data %>% filter(pathway=='visual'),"deprel",MINCOUNT)
ggplot(subset, aes(x=deprel, y=score)) + 
  geom_boxplot(outlier.colour = 'grey', outlier.size = 1, varwidth = FALSE) +
  coord_flip() +
  theme(text=element_text(size=15))
ggsave(file="visual-omission-deprel-boxplot.png",width=5, height=5)

subset <- filter_by_count(data %>% filter(pathway=='textual'),"pos",MINCOUNT)
ggplot(subset, aes(x=pos, y=score)) + 
  geom_boxplot(outlier.colour = 'grey', outlier.size = 1, varwidth = FALSE) +
  coord_flip() +
  theme(text=element_text(size=15))
ggsave(file="textual-omission-pos-boxplot.png",
       width=5, height=5)

subset <- filter_by_count(data %>% filter(pathway=='textual'),"deprel",MINCOUNT)
ggplot(subset, aes(x=deprel, y=score)) + 
  geom_boxplot(outlier.colour = 'grey', outlier.size = 1, varwidth = FALSE) +
  coord_flip() +
  theme(text=element_text(size=15))
ggsave(file="textual-omission-deprel-boxplot.png",
       width=5, height=5)

subset <- filter_by_count(data,"pos",MINCOUNT*2)
ggplot(subset, aes(x=pos, y=score, color=pathway)) + 
  geom_boxplot(outlier.size = 1, outlier.colour='grey', varwidth = FALSE) +
  coord_flip() +
  theme(text=element_text(size=15))
ggsave(file="imaginet-omission-pos-boxplot.png",
       width=5, height=5)

subset <- filter_by_count(data,"deprel",MINCOUNT*2)
ggplot(subset, aes(x=deprel, y=score, color=pathway)) + 
  geom_boxplot(outlier.size = 1, outlier.colour='grey', varwidth = FALSE) +
  coord_flip() +
  theme(text=element_text(size=15))
ggsave(file="imaginet-omission-deprel-boxplot.png",
       width=5, height=5)

```
## Skipthought
```{r}
#data_skip <- rbind(data_v, read.data("../data//skipthought_omission.csv") %>% mutate(pathway="skipthought")) %>% rename(model=pathway)

#ggplot(data_skip %>% filter_by_count("deprel",MINCOUNT*2), aes(x=deprel, y=score, color=model)) +
#  geom_boxplot(outlier.size=1, varwidth=FALSE) + 
#   coord_flip() +
#   theme(text=element_text(size=15))
# ggsave(file="skipthought-omission-deprel-boxplot.png",
#        width=5, height=5)
```

### Log scores ratios per dependency
```{r}
ratio <- filter(data, pathway=='visual')$score / filter(data, pathway=='textual')$score  
data_r <- filter(data, pathway=='visual')
data_r$ratio <- ratio
data_r$deprel <- reorder(data_r$deprel, data_r$ratio, median)
data_r$pos <- reorder(data_r$pos, data_r$ratio, median)
```
```{r}
ggplot(filter_by_count(data_r,"pos",MINCOUNT) %>% filter, 
       aes(x=pos, y=log(ratio))) + 
  geom_boxplot(outlier.colour = 'grey', outlier.size = 1, varwidth = FALSE) + 
  coord_flip() +
  theme(text=element_text(size=15))
ggsave(file="omission-pos-boxplot.png", width=5, height=5)
ggplot(filter_by_count(data_r,"deprel",MINCOUNT), aes(x=deprel, y=log(ratio))) + 
  geom_boxplot(outlier.colour = 'grey', outlier.size = 1, varwidth = FALSE) + 
  coord_flip() +
  theme(text=element_text(size=15))
ggsave(file="omission-deprel-boxplot.png", width=5, height=5)
m1 <- lm(log(ratio) ~ pos, data=data_r)
m2 <- lm(log(ratio) ~ deprel, data=data_r)
m3 <- lm(log(ratio) ~ deprel+pos, data=data_r)

summary(m1)
summary(m2)
summary(m3)
```
### Sentiment
```{r}
# data_sent <- read.data("../data/sentiment_omission.csv") 
# data_sent$deprel <- reorder(data_sent$deprel, data_sent$score, median)
# data_sent$pos <- reorder(data_sent$pos, data_sent$score, median)
# 
#             
# subset <- filter_by_count(data_sent,"pos",300)
# p.sent.pos <- ggplot(subset, aes(x=pos, y=score)) + 
#   geom_boxplot(outlier.colour = 'grey', outlier.size = 1, varwidth = FALSE) 
# ylim1 <- boxplot.stats(filter(subset, pos=="JJS")$score)$stats[c(1, 5)]
# p.sent.pos + 
#   coord_flip(ylim=ylim1*1.05) +
#   theme(text=element_text(size=15))
# ggsave(file="sent-omission-pos-boxplot.png",width=5, height=5)
# 
# subset <- filter_by_count(data_sent,"deprel",300)
# p.sent.deprel <- ggplot(subset, aes(x=deprel, y=score)) + 
#   geom_boxplot(outlier.colour = 'grey', outlier.size = 1, varwidth = FALSE) 
# ylim1 <- boxplot.stats(filter(subset, deprel=="root")$score)$stats[c(1, 5)]
# p.sent.deprel + 
#   coord_flip(ylim=ylim1*1.05) +
#   theme(text=element_text(size=15))
# ggsave(file="sent-omission-deprel-boxplot.png",
#        width=5, height=5)
# 
# ggplot(data_r %>% filter(pos=="VBZ") %>% filter_by_count(col = "deprel", 100), aes(x=deprel, y=log(ratio))) + geom_boxplot(varwidth = TRUE) + coord_flip()
# ggsave(file="omission-vbz-boxplot.pdf")
```

```{r}
data <- read.csv("visual_omission_with_position.csv")
data <- data %>% mutate(pos=reorder(pos, score),
                        deprel=reorder(deprel, score)) #%>% filter(pos %in% c("NN","NNS"))
#m0 <- lm(log(score) ~ pos, data=data)
#m1 <- lm(log(score) ~ pos + as.factor(position), data=data)
#summary(m0)
#summary(m1)
cor(data %>% select(position, score), method = "pearson")
cor(data  %>% filter(pos %in% c("NN","NNP","NNS")) %>% select(position, score), method = "pearson")

ggplot(data %>% filter_by_count("pos",MINCOUNT*2), 
            aes(x=position, color=pos, y=score)) +
  geom_point(position = "jitter", alpha=0.03) +
  geom_smooth() +
  xlim(0,20) +
  ylim(0.0, 0.5) +
  theme(aspect.ratio=2/3, text=element_text(size=25)) + 
  ggtitle("Visual")
ggsave(file="visual-omission-by-position.png")

```

```{r}
data <- read.csv("textual_omission_with_position.csv")
data <- data %>% mutate(pos=reorder(pos, score),
                        deprel=reorder(deprel, score))
cor(data  %>% select(position, score), method = "pearson")
#m0 <- lm(log(score) ~ pos, data=data)
#m1 <- lm(log(score) ~ pos + as.factor(position), data=data)
#summary(m0)
#summary(m1)

ggplot(data %>% filter_by_count("pos",MINCOUNT*2), 
            aes(x=position, color=pos, y=score)) +
  geom_point(position = "jitter", alpha=0.03) +
  geom_smooth() +
  xlim(0,20) +
  ylim(0.0, 0.5) + 
  theme(aspect.ratio=2/3, text=element_text(size=25)) +
  ggtitle("Textual")
ggsave(file="textual-omission-by-position.png")

```
```{r}
library(dplyr)
library(ggplot2)

top_words <- read.csv("top_words.csv") %>% 
  mutate(deprel = reorder(deprel, score))
ggplot(top_words %>% filter_by_count("deprel",70), 
       aes(x=word, y=score, color=deprel)) +
  geom_boxplot(position = "dodge") +
   theme(text=element_text(size=25))
ggsave(file="top_words.pdf")
```