### 2017-05-16 doc/original_reviews.tex

\documentclass[10pt,a4paper]{letter}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{url}

\begin{document}

I'm writing to you in my capacity as the TACL action editor for your
submission 882, "Representation of linguistic form and function in recurrent
neural networks". While we are not accepting your paper in its current form,
we do encourage you to revise and resubmit it within 3-6 months.

We received three reports. These are included below. Please do consider all
of the points they make as you revise the work. Some high-level themes:

* The paper is pitched in very broad terms (methods for analyzing RNNs), but
the methods are actually quite specific to classifier-type RNNs, and many
aspects of the work seem applicable only to the Imaginet setting. I would
suggest clarifying the scope of the claims.

* Similarly, as Reviewer A points out, some of the rhetoric is misaligned
with the methods and results.

* The core results for content words and functions words seem like the
expected ones given their linguistic content. In particular, the omission
score method seems to show that the networks learn to identify
discriminating content words, and that
changing or removing those words has a dramatic effect on final predictions.
This seems like a prerequisite (given the linguistic facts) for learning an
effective classifier. Regular train/dev/test experimental analyses already
show that these models are effective. Could you further articulate what we
learn via the omission score method that we don't get from standard
evaluations?

* The connections with POS tags and dependency relations are intriguing.
Could you, though, clarify the significance of these findings? My concern is
that they just reflect the fact that the content-word/function-word
distinction is highly correlated with POS and dependency relation.

* I had trouble understanding the information structure section. I agree
with the core insight about old/topic before new/comment in languages like
English. I had trouble relating this to what I would expect from the RNN.
For instance, the structure of the RNN seems to ensure that information
nearer the end of the sequence will have a larger impact. Is the Textual
omission score result due to this factor or is it truly due to something
about language? Conversely, for Visual, I guess I would have anticipated
exactly the opposite of what is reported here: the comment part should be
the most important, corresponding to the new and visually salient elements,
rather than those that can fade into the background.

For additional questions and concerns, and some suggestions for how to
improve accessibility and readability, please see the full reports.

If you do choose to revise and resubmit, please make use a *new* submission
number, and follow the instructions in section "Revision and Resubmission
Policy for TACL Submissions" at
\url{https://tacl2013.cs.columbia.edu/ojs/index.php/tacl/about/submissions#authorGuidelines.}
We  allow you 1-2 additional pages in the revised version for addressing the
referees' concerns.

Please understand that while we have endeavored to provide some guidance on
how to revise the manuscript, the paper will be reviewed afresh should you
choose to resubmit (possibly involving a change of action editor and
reviewers), with no guarantee of acceptance.

Thank you for considering TACL for your work, and again, I encourage you to
revise and resubmit within the specified timeframe.

---Chris

Christopher Potts
Stanford University
cgpotts@stanford.edu

-----------------------
THE REVIEWS
----------------------- \\
Reviewer A:

CLARITY: For the reasonably well-prepared reader, is it clear what was done
and why? Is the paper well-written and well-structured?:
        3. Mostly understandable to me with some effort.

ORIGINALITY/INNOVATIVENESS: How original is the approach? Does this paper
break new ground in topic, methodology, or content? How exciting and
innovative is the research it describes?
Note that a paper could score high for originality even if the results do
not show a convincing benefit.
:
        2. Pedestrian: Obvious, or a minor improvement on familiar techniques.

SOUNDNESS/CORRECTNESS: First, is the technical approach sound and
well-chosen? Second, can one trust the claims of the paper -- are they
supported by properexperiments and are the results of the experiments
correctlyinterpreted?:
        3. Fairly reasonable work. The approach is not bad, and at least the main
claims are probably correct, but I am not entirely ready to accept them
(based on the material in the paper).

MEANINGFUL COMPARISON: Does the author make clear where the presented system
sits with respect to existing literature? Are the references adequate?:
        3. Bibliography and comparison are somewhat helpful, but it could be hard
for a reader to determine exactly how this work relates to previous work or
what its benefits and limitations are.

SUBSTANCE: Does this paper have enough substance (in terms of the amount of
work), or would it benefit from more ideas or analysis?:
        2. Work in progress. There are enough good ideas, but perhaps not enough
results yet.

IMPACT OF IDEAS OR RESULTS: How significant is the work described? If the
ideas are novel, will they also be useful or inspirational? If the results
are sound, are they also important? Does the paper bring new insights into
the nature of the problem?:
        3. Interesting but not too influential. The work will be cited, but mainly
for comparison or as a source of minor contributions.

REPLICABILITY: Will members of the ACL community be able to reproduce or
verify the results in this paper?:
        3. could reproduce the results with some difficulty. The settings of
parameters are underspecified or subjectively determined; the
training/evaluation data are not widely available.

IMPACT OF PROMISED SOFTWARE:  If the authors state (in anonymous fashion)
that their software will be available, what is the expected impact of the
software package?:
        1. No usable software released.

IMPACT OF PROMISED DATASET(S): If the authors state (in anonymous fashion)
that datasets will be released, how valuable will they be to others?:
        1. No usable datasets submitted.

TACL-WORTHY AS IS? In answering, think over all your scores above. If a
paper has some weaknesses, but you really got a lot out of it, feel free to
recommend it. If a paper is solid but you could live without it, let us know
that you're ambivalent.

Note: after you submit this review form, you'll need to answer a related but
different question via a pull-down menu: how long would it take for the
authors to revise the submission to be TACL-worthy?
:
        3. Ambivalent: OK but does not seem up to the standards of TACL.

Detailed Comments for the Authors:
        The paper presents some empirical studies on the task of visualising the
internal representations of the RNN on some language tasks. The authors
specifically focused on the RNN architectures proposed by Chrupala et al
(2015) and focused on two tasks - language modelling and predicting visual
representations from word sequences.

The authors focused on the discussions on the understanding of the hidden
activation vectors learned at the end of RNN as well as the understanding of
the dimensions in the learned vector representations. For the first task,
the authors used the "omission" trick -- excluding a certain single word in
the sentence, and generate a different vector representation at the end, and
make comparisons with the original vector representation. The authors then
tried to draw some conclusions on the importance of each word based on the
omission scores.

One question about the omission score: the lower the score is, the low the
similarity is. Thus, when dropping the important words, the similarity
should go down -- so a lower omission score. It is not clear to me why
dropping important words will lead to a higher omission score.

Although I agree it's interesting to do such analysis: especially here we
have two different tasks -- one involves visual aspects which is very nice,
there are two important questions. First, why is the architecture of
Chrupala et al (2015) selected only? Is it representative enough for general
RNNs? What about LSTM and its variants for language modelling task? How
would they be different? It would be good to see such discussions. Second,
most of the paper presents qualitative results, without rigorous
quantitative justifications. It would be nice, however, if the authors could
collect some manual annotations and conduct automatic evaluations based on
such annotations. For example, such annotations could include human's
judgements on how important each word (or word class, syntactic class) in
the sentence are etc. Finally, although the visualisation is helpful and it
might help the readers understand the underlying activities of the RNN,
little is known about why. I understand it is just an empirical paper,
however, it would be helpful to at least share with the readers some high
level insights so that the readers could learn more from this paper. One
question that I have in mind after reading this paper is: these results
related to RNN presented here are expected. It's good that RNN is able to
capture such things. However, how do I make use of the results here in my
research? How does this have some impact on my work? The authors claimed at
the end of the paper that "The goal of our paper is to propose novel methods
for the encoding of linguistic knowledge in RNNs" -- I am not sure if I am
convinced that this is goal of this work. It is largely an empirical paper
that tries to visualise/understand the RNN which was previously regarded as
a blackbox, but does not really "propose" methods for encoding linguistic
knowledge in RNN.

The explanations of several places, such as Sec 6.1 and Sec 6.2 could be
improved. It was a little difficult for me to follow.

Minor: diffeent -> different.

REVIEWER CONFIDENCE:
        3. Pretty sure, but there's a chance I missed something. Although I have a
good feel for this area in general, I did not carefully check the paper's
details, e.g., the math, experimental design, or novelty.

-----------------------
Reviewer B:
-----------------------

CLARITY: For the reasonably well-prepared reader, is it clear what was done
and why? Is the paper well-written and well-structured?:
        5. Very clear.

ORIGINALITY/INNOVATIVENESS: How original is the approach? Does this paper
break new ground in topic, methodology, or content? How exciting and
innovative is the research it describes?
Note that a paper could score high for originality even if the results do
not show a convincing benefit.
:
        3. Respectable: A nice research contribution that represents a notable
extension of prior approaches or methodologies.

SOUNDNESS/CORRECTNESS: First, is the technical approach sound and
well-chosen? Second, can one trust the claims of the paper -- are they
supported by properexperiments and are the results of the experiments
correctlyinterpreted?:
        4. Generally solid work, although there are some aspects of the approach or
evaluation I am not sure about.

MEANINGFUL COMPARISON: Does the author make clear where the presented system
sits with respect to existing literature? Are the references adequate?:
        4. Mostly solid bibliography and comparison, but there are a few additional
references that should be included. Discussion of benefits and limitations
is acceptable but not enlightening.

SUBSTANCE: Does this paper have enough substance (in terms of the amount of
work), or would it benefit from more ideas or analysis?:
        4. Represents an appropriate amount of work for a publication in this
journal. (most submissions)

IMPACT OF IDEAS OR RESULTS: How significant is the work described? If the
ideas are novel, will they also be useful or inspirational? If the results
are sound, are they also important? Does the paper bring new insights into
the nature of the problem?:
        4. Some of the ideas or results will substantially help other people's
ongoing research.

REPLICABILITY: Will members of the ACL community be able to reproduce or
verify the results in this paper?:
        4. could mostly reproduce the results, but there may be some
variation because of sample variance or minor variations in their
interpretation of the protocol or method.

IMPACT OF PROMISED SOFTWARE:  If the authors state (in anonymous fashion)
that their software will be available, what is the expected impact of the
software package?:
        1. No usable software released.

IMPACT OF PROMISED DATASET(S): If the authors state (in anonymous fashion)
that datasets will be released, how valuable will they be to others?:
        1. No usable datasets submitted.

TACL-WORTHY AS IS? In answering, think over all your scores above. If a
paper has some weaknesses, but you really got a lot out of it, feel free to
recommend it. If a paper is solid but you could live without it, let us know
that you're ambivalent.

Note: after you submit this review form, you'll need to answer a related but
different question via a pull-down menu: how long would it take for the
authors to revise the submission to be TACL-worthy?
:
        4. Worthy: A good paper that is worthy of being published in TACL.

Detailed Comments for the Authors:
        This paper provides an in-depth qualitative and quantitative of a joint
caption-generation and image selection model (IMAGINET). It casts the said
analysis as a general procedure for introspecting a trained model through
several methods. It examines the effect of removing words on the final
representation of a text-predicting RNN and an image predicting RNN as a
measure of word, POS, dependency relation, and positional importance. It
also examines how various words and word classes contribute to increases in
hidden layer activation, or preservation thereof. The results are entirely
unsurprising, in that content words matter for image matching, and function
words affect textual meaning while preserving information in the hidden
layer. The fact that these results match obvious intuitions should not be
taken as criticism of the paper in any way, and I am satisfied that the
method by which they were obtained is sound.

Overall, the paper is clear, well written, well presented. Such attempts to
analyse trained recurrent neural networks from a linguistic perspective is
furthermore a timely and welcome contribution, given their near-ubiquity in
*ACL publications (at least in semantics and ML tracks). The results of the
analysis are convincing, and I would support publication of this paper in
TACL. Below are a few criticisms and suggested improvements that would help
improve the paper.

The main criticism I would have of this paper is that it sets out as a
providing generic new method for analysing RNNs with text input, amongst
other modalities. I think this is overstating the contents of the paper: the
methods of analysis are fairly tied to the nature of the model
(unidirectional RNN without attention) and broad task (producing a final
representation that can be used for classification or generation). It would
have been interesting to the same analyses performed on other, related
models (bi-directional RNN, word-based log-bilinear models, etc), and on RNN
tasks with other objectives (machine translation, language modelling,
sequence labelling). I realise it's a tall order to add this to the current
work, but perhaps re-stating the scope of the work (providing an example of
how to perform further qualitative and quantitative analysis of a an
existing model trained on a specific task) and briefly discussing how to
adapt the experiments to other domains would be good. Likewise, a discussion
comparing and contrasting these methods of analysis with other
(complementary?) visualisation methods for RNNs with attention such at
Badhanau et al. 2015 or Rocktaschel et al. 2015 [ICLR 2016] (which could be
cited here) would be welcome. To summarise this point, I do not believe the
authors have willingly misrepresented the aim and scope of their work, but
think it would benefit from being restated a bit more clearly, bearing in
mind the comments made above.

Other more minor comments:

* It would be nice to get a clear statement of the objective being optimised
against in eq 9.
* Figure 2 a little hard to read.
* Figure 6 very hard to read.
* Section 5.1 "textual is more sensitive to ..." -> it looks like textual is
more sensitive to most words being removed to some extent. I don't know if
it's fair to single out these categories. Instead, just contrast more
clearly with the heavy focus on content words (mainly NNs) for Visual.
* Section 5.2, model 1 vs model 2 <- is the a significant difference in
number of features here? Model 2 seems like it would have more features, and
intuitively, more info == better model (modulo feature quality), so quick
comment on this here would be good.
* Section 5.3 is very tied to structure of the model use in this paper
(would expect different results for bidirection RNN, recursive neural net).
Comment on this here would be good.
* Not a fan of qualitative analysis as done in Table 1/2. Too easy
over-interpret, but I realise this is personal taste. Would not be offended
if this is left in, but would suggest scrapping this to create space for
discussions suggested above.

REVIEWER CONFIDENCE:
        5. Positive that my evaluation is correct. I read the paper very carefully
and am familiar with related work.

-----------------------
Reviewer C:
-----------------------

CLARITY: For the reasonably well-prepared reader, is it clear what was done
and why? Is the paper well-written and well-structured?:
        3. Mostly understandable to me with some effort.

ORIGINALITY/INNOVATIVENESS: How original is the approach? Does this paper
break new ground in topic, methodology, or content? How exciting and
innovative is the research it describes?
Note that a paper could score high for originality even if the results do
not show a convincing benefit.
:
        4. Creative: An intriguing problem, technique, or approach that is
substantially different from previous research.

SOUNDNESS/CORRECTNESS: First, is the technical approach sound and
well-chosen? Second, can one trust the claims of the paper -- are they
supported by properexperiments and are the results of the experiments
correctlyinterpreted?:
        4. Generally solid work, although there are some aspects of the approach or
evaluation I am not sure about.

MEANINGFUL COMPARISON: Does the author make clear where the presented system
sits with respect to existing literature? Are the references adequate?:
        4. Mostly solid bibliography and comparison, but there are a few additional
references that should be included. Discussion of benefits and limitations
is acceptable but not enlightening.

SUBSTANCE: Does this paper have enough substance (in terms of the amount of
work), or would it benefit from more ideas or analysis?:
        4. Represents an appropriate amount of work for a publication in this
journal. (most submissions)

IMPACT OF IDEAS OR RESULTS: How significant is the work described? If the
ideas are novel, will they also be useful or inspirational? If the results
are sound, are they also important? Does the paper bring new insights into
the nature of the problem?:
        5. Will affect the field by altering other people's choice of research
topics or basic approach.

REPLICABILITY: Will members of the ACL community be able to reproduce or
verify the results in this paper?:
        3. could reproduce the results with some difficulty. The settings of
parameters are underspecified or subjectively determined; the
training/evaluation data are not widely available.

IMPACT OF PROMISED SOFTWARE:  If the authors state (in anonymous fashion)
that their software will be available, what is the expected impact of the
software package?:
        1. No usable software released.

IMPACT OF PROMISED DATASET(S): If the authors state (in anonymous fashion)
that datasets will be released, how valuable will they be to others?:
        4. Useful: I would recommend the new datasets to other researchers or
developers for their ongoing work.

TACL-WORTHY AS IS? In answering, think over all your scores above. If a
paper has some weaknesses, but you really got a lot out of it, feel free to
recommend it. If a paper is solid but you could live without it, let us know
that you're ambivalent.

Note: after you submit this review form, you'll need to answer a related but
different question via a pull-down menu: how long would it take for the
authors to revise the submission to be TACL-worthy?
:
        5. Strong: I'd like to see it accepted; it will be one of the better papers
in TACL.

Detailed Comments for the Authors:
        Summary :

This paper addresses questions on visualization and understanding of
activation patterns of recurrent neural networks given a paired
linguistic-visual representational input. Though sharing the same word
embeddings, the visual and linguistic neural pathways triggered different
activation patterns, manifested in the form of different omission scores
distributions at the end of each pathway, when variants of the same input
sentence are given as inputs to the RNN. The paper also gave an in-depth
analysis of the hidden units by comparative evaluations of "contexts", which
are encodings at the level of the hidden units, between the two neural
pathways.

Merits:

- novel visualization into inner workings of NN, and from an
easier-to-grasp, higher level linguistic perspective. To my knowledge, NN
are known to outperform many other learners, yet efforts on
visualization/understanding of their inner layers is rare due to their
inherent complexity.

- this work has the ability to generalize from dual linguistics
representations to n-modal representation, where there are increasingly more
work on multimodal-feature approach to solving old tasks. The principled
approach used in this paper can be used to explain away the orthogonal
contribution provided by each neural pathway trigged by each of the
modalities to gain a better understanding of the network.

Issues/concerns:

The paper builds heavily on previous work "IMAGINET", is densely described
but lacking details in certain sections. For example, I find it rather
challenging to understand the following:

- the process of image vector prediction is vague (eqn 7). Does it mean
prediction at the pixel level of an entire image, or a blob or principal
image segment of interest? Also, how does the image retrieval shown in
Figure 1 operate?

- Figure 6 would benefit from showing the aggregated POS groupings such as
"VBN,VBG,VB" all grouped into "VERB" etc. for benefit of clarity in a small
graph (too cluttered)

REVIEWER CONFIDENCE:
        3. Pretty sure, but there's a chance I missed something. Although I have a
good feel for this area in general, I did not carefully check the paper's
details, e.g., the math, experimental design, or novelty.

\end{document}