### 2017-05-16 doc/regression.tex

\section{Representing syntactic information}
\label{sec:regression}

It is assumed that to perform a task that involves some level of language 
understanding, RNNs need to learn some level syntactic knowledge from the 
linguistic input. To uncover if the networks indeed represent syntactic 
information, we trained a number of Logistic Regression models on various 
combinations of n-gram tokens in the input as well as the hidden states of the 
networks to predict the syntactic categories of the tokens at each time-step $t$. 
From the sentences we extracted unigram and n-gram features up to a window 
size of 4; for example, to predict the label for {\it dog} in the sentence {\it the nice 
dog} we extract {\it the}$_2$, {\it nice}$_1$, {\it dog}$_0$, {\it the$_2$ nice$_1$}, 
{\it nice$_1$ dog$_0$}, and {\it the$_2$ nice$_1$ dog$_0$}, etc.

We trained several models on a combination of the following features:
\begin{itemize}
  \item current word $w_t$
  \item current word-embedding of {\sc Imaginet} $e_t^I$ (same for {\sc Visual} 
and {\sc Textual})
  \item current word-embedding of {\sc Sentiment} $e_t^S$
  \item 4-gram features
  \item hidden state of {\sc Visual} $h_t^V$
  \item hidden state of {\sc Textual} $h_t^T$
  \item hidden state of {\sc Sentiment} $h_t^S$
%  \item hidden state of {\sc Visual} $h_t^V$ + 4-gram features
%  \item hidden state of {\sc Textual} $h_t^T$ + 4-gram features
\end{itemize}

The motivation behind this experiment is that if the networks generalize over 
syntactic categories, then using their activation vectors as features will lead to a 
more accurate model f$h_t$or syntactic category prediction than only using the tokens 
themselves. If the embeddings $e_t$ and hidden activations $h_t$ outperform 
the word form $w_t$ and 4-gram models respectively, then the linguistic 
representations of the model do generalize over syntactic categories.

\begin{table}[]
\centering
\caption{Results of predicting syntactic categories for each token in the input to 
SENTIMENT and IMAGINET models, by Logistic Regression classifiers trained 
on features in the first column. The second and third columns report the micro F-
scores on DepRel and POS categories respectively.}

\label{tab:logistic}
\begin{tabular}{lll}
\hline
 \multicolumn{3}{c}{\sc Sentiment}           \\
 \hline 
{\bf Features}      &  {\bf deprel} &  {\bf POS}     \\
\hline 
word $w_t$                   		  & .50  & .65   \\
4-gram                       		  & .62  & .71   \\
$e^S_{t}$          & .33  & .38   \\ 
$h_t^S$                      		  & .30  & .32    \\
\hline 
\hline
\multicolumn{3}{c}{\sc Imaginet}                          \\
 \hline 
{\bf Features}      &  {\bf deprel} &  {\bf POS}     \\
\hline 
$w_t$                   		  	      & .710   & .93  \\
4-gram                         		  & .807  & .93  \\
$e^I_{t}$          			  	      & .73   & .95  \\ 
$h_t^V$                      		  & .751  & .88  \\
$h_t^T$                      		  & .811  & .94  \\
$h_t^T$ + $h_t^V$            		  & .81   & .94  \\
$h_t^V$ + 4-gram             		  & .809   & .93  \\
$h_t^T$ + 4-gram             		  & .823   & .95
\end{tabular}
\end{table}

Table~\ref{tab:logistic} shows the results of predicting the syntactic categories of 
tokens in the input sentences using different (combinations of) features. 
\todo{Are we going to change this to test set?}

The results for {\sc Sentiment} show that the representations (both 
$e^S_{t}$ and $h_t^S$) are less predictive of the grammatical categories of the 
target words than using the unigram and 4-gram features.

%In fact there are only 8 categories out of 44 with non-zero f-scores for deprels: 
%amod (0.28), cc (.64), conj (.13), det (.56), nsubj (.29), pobj (.30), prep (.43) and 
%root (.26).
However, in case of {\sc Imaginet}, using the embeddings $e_t$ leads to higher F-score for both dependency category and part-of-speech tag prediction of the tokens $w_t$ than using $w_{t}$'s 
themselves. This result provides evidence that syntactic information is indeed encoded in the learned word embeddings. 
Similarly, the classifier trained on the hidden activations of {\sc Textual } $h_t^T$ 
performs better on both DepRel and POS prediction than using 4-grams features, but only by a 
narrow margin.

The more interesting results, however, are the low F-scores when 
using $h_{t}^V$ as features. These results suggest that {\sc Visual} represents less 
syntactic information then {\sc Textual} about sequences of words. We also observe 
that the maximum F-score is achieved using the combinations of $h_{t}^T$ and the 4-
gram features, suggesting that perhaps some information about the word identities is 
lost in $h_{t}^T$.