### 2017-05-29 doc/experiments.tex

\section{Experiments}
\label{sec:experiments}

In the following sections, we report a series of experiments in which
we explore the kinds of linguistic regularities the
networks learn from
word-level input. Section~\ref{sec:macro} presents the macro-level
analyses where we propose methods to analyze the final hidden activation 
vectors of the recurrent pathways from linguistic point of view. 
Section~\ref{sec:micro} reports exploratory experiments on
the linguistic features encoded by individual hidden units. 
In both sections we report our findings based on the {\sc Imaginet}
model, discuss the generalizabilty of our methods to other architectures.
\label{edit:experimentsgeneral}, and whenever appropriate compare to
our two other models, {\sc LM} and {\sc Sum}.  For all the
experiments, we trained {\sc Imaginet} on the training portion of the
MSCOCO image-caption dataset \cite{lin2014microsoft}, and analyzed the
representations of the sentences in the validation set corresponding
to 5000 randomly chosen images. The target image representations were
extracted from the pre-softmax layer of the 16-layer CNN of
\namecite{simonyan2014very}.

% \iffalse

% {\bf Section~\ref{sec:macro}} includes experiments that provide linguistic analysis of the activation patterns on the activation vector level with results more quantitative in nature. Section~\ref{sec:salience} describes a method to estimate the salience of tokens as a function of their part-of-speech category and grammatical function (dependency relation) in sentences.

%  Sections~\ref{sec:gramfunc} and \ref{subsec:information-struct} apply this method and provide evidence that the {\sc Visual} pathway of {\sc Imaginet} learns to interpret the same word fullfilling different grammatical functions differently, and is sensitive to the information structure of sentences.

% {\bf Section~\ref{sec:micro}} compliments the macro level analysis and reports a series of more qualitative experiments exploring the kinds of linguistic abstraction represented by individual hidden units. Section~\ref{sec:reprdim} proposes a simple method for representing
% the function of hidden units by identifying the top K contexts that yield the highest activation values, and Section~\ref{sec:topk} provides a general qualitative analysis of these contexts.

% For the experiments using the {\sc Imaginet} model we use the validation set of MS-COCO.
% Finally Section~\ref{sec:syntacticdim} goes more into detail and shows examples 
% of hidden units that are active for particular multi-word constructions with joint
% semantical and syntactic regularities.  
% \fi

%\begin{enumerate}
%\item How accurately does each model represent syntactic knowledge? WE DONT DO THIS
%\item What types of grammatical functions each model pays most attention to? WE DO THIS
%\item What is the level of separation between syntactic and semantic 
%representations learned by the models? NO CHANCE
%\item What functions are the individual units in each network specialized for? WE DO THIS 
%\item On each of the above dimensions, what are the systematic differences 
%between models that are optimized for different tasks?  WE ALSO DO THIS
%\end{enumerate}

%Section~\ref{sec:categories} looks at the salience of two sets of syntactic categories in 
%each model, by estimating how much input tokens from each category influence 
%the final performance in each task. In Section~\ref{sec:syntax}, we examine the 
%degree to which syntactic information is encoded in the internal representations 
%acquired by each model. Section~\ref{sec:dimension} focuses on analyzing the 
%reaction patterns of individual units in each model, and on finding out specialized 
%functions that these units perform.

%Finally in Section~\ref{sec:visualize}, we look at 
%how the meaning representations of word sequences evolve over time for each 
%model. \todo{not sure how to relate this last one to our objectives.}
%In all our experiments, we compare the nature and degree of the acquired 
%syntactic knowledge among models that are trained for performing different tasks.

%\noindent{\bf Syntactic categories.} In all the following analyses, we use two different sets of syntactic categories: part-of-speech categories (POS) and dependency relations (DepRel). The tagging is performed jointly using the TurboParser dependency parser\footnote{This implementation is available at https://github.com/andre-martins/TurboParser.} \cite{martins2013turning}. For POS tags we use the Penn Treebank tagset and for the dependency relations the Stanford basic dependencies.

%\input{regression}
\input{categories}
\input{dimensions}
%\input{visualization}