### 2017-07-12 doc/cover.tex
adding style files and cover letter

\documentclass{letter}
\signature{Akos Kadar \\ Grzegorz Chrupala \\ Afra Alishahi}

\begin{document}
\begin{letter}{}
\opening{Dear Dr.~Merlo,}

Thank you for informing us about the outcome of the review process, we are very pleased that the paper has been accepted to be published in Computational Linguistics.

Attached please find the final version of the article. We have applied the formatting requirements as specified in the author guidelines, and updated the references. Please let us know if any other information is required.

\closing{Sincerely,}

\end{letter}
\end{document}