### 2017-07-17 doc/double_paper.tex
Added double-spaced version of paper.

%\documentclass[shortpaper]{clv3}  %for single-spaced version
\documentclass[manuscript]{clv3} %for double-spaced version

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{algorithm}

\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{amssymb}
\usepackage{pifont}
\usepackage{todonotes}
\usepackage{url}
\usepackage{array,multirow,graphicx}
\usepackage{rotating}
\usepackage{times}
\usepackage{latexsym}
\usepackage{arydshln}
\usepackage{enumitem}
\setlist{nolistsep} % needs enumitem

\issue{43}{3}{2017}

\title{Representation of linguistic form and function in recurrent neural networks}
\runningtitle{Representation of linguistic form and function in recurrent neural networks}

\author{Ákos Kádár \thanks{Tilburg Center for Cognition and Communication,
Tilburg University, 5000 LE Tilburg, The Netherlands, 
E-mail: \texttt{\{a.kadar, g.chrupala, a.alishahi\}@uvt.nl}} }  
\affil{Tilburg University}

\author{Grzegorz Chrupała \footnotemark[1]}
\affil{Tilburg University}

\author{Afra Alishahi \footnotemark[1]}
\affil{Tilburg University}

\runningauthor{Kádár, Chrupała and Alishahi}

\historydates{Submission received: 21st July, 2016 \\
Revised version received: 5th June, 2017 \\
Accepted for publication: 7th July, 2017}

\begin{document}
\maketitle

\input{intro}
\input{related}
\input{models}
\input{experiments}
\input{conclusions}

\bibliographystyle{compling}
%\starttwocolumn http://cljournal.org/commonprobs.html says we need
%this, but the template doesn't do it
\bibliography{refs}

\end{document}